Designing Modular Software: A Case Study in Introductory Statistics

### 2016-09-29 jcgs-paper-2016/paper.Rmd
Reproducibility section

>>>>

While web-based tools written using Shiny (including `intRo`) have appealing characteristics such as being multi-platform, requiring no installation, and requiring no software licenses, one limitation immediately presents itself. The actions taken in the application are typically not reproducible as in a standard R script. We have designed `intRo` to overcome this limitation by capturing the unevaluated expression of all actions taken by the user in the interface. This expression is then parsed and printed in a code window at the bottom, while simultaneously being executed by the R process running on the server.

In essence, this procedure transcribes user actions in a GUI to R code. When run in a standard R session, the results produced will be identical to the results shown in `intRo`. The full series of actions taken by the user are transcribed and can then be exchanged by researchers, students, and developers in a manner similar to normal scripting. Even "printing" the results of an `intRo` session amounts to nothing more than executing the given code on the server, and then storing the results in an rmarkdown document, weaving the code with the results to produce a full compendium. While not strictly necessary, this lends credibility to the results produced by `intRo` in the sense that they are directly reproduced by the server every time the user clicks the print button.

Reproducibility in `intRo` is accomplished with the previously mentioned `interpolate` function. `interpolate` accepts an expression and an arbitrary number of arguments as an argument, substitutes the arguments into the expression, prints the results to the console, and evaluates the parsed expression. This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`. One potential drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).

One potential enhancement to this framework would be the inclusion of state-saving and state-resuming. Because an `intRo` session is uniquely represented by the series of commands stored as code, the code itself could represent a checkpoint for resuming a new `intRo` session. Currently, beginning a new session will start the application with no memory of previous sessions. In real-world usage, state saving could allow a user to continue work later. At this time, this can only be done by taking the code and pasting it into a standard R session, although such an enhancement would likely involve minimal changes to `intRo`'s underlying structure.s
